---
title: "p8105_hw2_lk2776"
output: github_document
date: "2024-09-30"
---
## hw2

```{r setup, warning=FALSE, message=FALSE}
#load packages; print 3 lines of tibble output
library(tidyverse)
options(tibble.print_min=5)
```

problem1:

The cleaned NYC dataset has 19 columns and 1868 rows. The cleaning process involved reading the data with missing value using na(), cleaning the column names using janitor::clean_names(), selecting the desired columns using select(), and converting entry variable using mutate() and ifelse() functions. 

The dataset has line, station_name, station_latitude, station_longitude, routes1-11, entry, vending, entrance_type, ada variables. 


```{r study data}
#read data using read_csv() from tidyverse
nyc_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
                    na=c("NA",".",""), 
                    col_names=TRUE) |>
  #clean column names
  janitor::clean_names() |> 
  #select columns 
  select(line,station_name,station_latitude,station_longitude,
         route1,route2,route3,route4,route5,
         route6,route7,route8,route9,route10,route11,
         entry,vending,entrance_type,ada
         ) |> 
  #convert entry column from character to logical vector
  mutate(entry=if_else(entry %in%"YES",TRUE,FALSE))
  
head(nyc_data,5)
#check dimensions of dataset
dim_desc(nyc_data)
#check class of columns of dataset
nyc_data |> sapply(class)
```

