---
title: "p8105_hw2_lk2776"
output: github_document
date: "2024-09-30"
---
## hw2

```{r setup, warning=FALSE, message=FALSE}
#load packages; print 3 lines of tibble output
library(tidyverse)
options(tibble.print_min=5)
```

problem1:

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?

The cleaned NYC dataset has 19 columns and 1868 rows. The cleaning process involved reading the data with missing value using na(), cleaning the column names using janitor::clean_names(), selecting the desired columns using select(), and converting entry variable using mutate() and ifelse() functions. 

The dataset has line, station_name, station_latitude, station_longitude, routes1-11, entry, vending, entrance_type, ada variables. 


```{r study data}
#read data using read_csv() from tidyverse
nyc_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
                    na=c("NA",".",""), 
                    col_names=TRUE) |>
  #clean column names
  janitor::clean_names() |> 
  #select columns 
  select(line,station_name,station_latitude,station_longitude,
         entry,vending,entrance_type,ada,
         route1,route2,route3,route4,route5,
         route6,route7,route8,route9,route10,route11
         ) |> 
  #convert entry column from character to logical vector
  mutate(entry=if_else(entry %in%"YES",TRUE,FALSE))

#check class of columns of dataset to verify entry varible class
nyc_data |> sapply(class)

#check dimensions of dataset
dim_desc(nyc_data)

```

```{r answering specific questions}
#number of distinct line_station_names
nyc_data |> select(line,station_name)  |>
  unite("line_station_name", line:station_name) |> 
 n_distinct() 
   
#number of stations that are ada compliant
table(nyc_data$ada)

#proportion of stations with entry but with out vending

```

```{r}

```

filter()